---
title: "p8105_hw1_fx2212"
author: "Fenglin Xie"
date: "2025-09-15"
output: github_document
---
# Problem 1
## The introduction of the dataset
```{r}
library(moderndive)
data("early_january_weather")
```
This dataset has `r ncol(early_january_weather)` variables and `r nrow(early_january_weather)` observations, including varibles as follows:`r names(early_january_weather)`. Besides, the mean temperature is `r mean(early_january_weather$temp)`.

## Make a scatterplot and the descriptions for the plot
```{r}
library(tidyverse)
temp_vs_time_plot <- ggplot(early_january_weather, 
                            aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time in Early January",
       x = "Time",
       y = "Temperature (°F)",
       color = "Humidity (%)") +
  theme_minimal()

print(temp_vs_time_plot)
```

From this plot, we can see the two points: 
+  Humidity levels show a relationship with temperature patterns, with higher humidity values often coinciding with periods of elevated temperature. 
+  There appears to be a slight overall increase in temperature values from January 7 to January 14

## Export the scatterplot to project directory
```{r}
ggsave("temp_vs_time_plot.png")
```

# Problem 2
## Create a data frame
```{r}
set.seed(123)

df <- tibble(
  norm_samp = rnorm(10, 0, 1),
  log_vec = norm_samp > 0,
  char_vec = letters[1:10],
  fact_vec = factor(rep(c("low", "medium", "high"), length.out = 10))
)

print(df)
```

## Calculate means for each variable
```{r}
mean_norm <- mean(pull(df, norm_samp))
mean_logical <- mean(pull(df, log_vec))
mean_character <- mean(pull(df, char_vec))
mean_factor <- mean(pull(df, fact_vec))
```

Only the numeric variable (norm_samp) and logical variable (log_vec) can be used to calculate a mean. The character variable (char_vec) and factor variable (fact_vec) cannot be used to calculate a mean, as attempting to do so would result in an error.

## Apply as.numeric to different variable types
```{r}
logical_to_numeric <- as.numeric(pull(df, log_vec))
character_to_numeric <- as.numeric(pull(df, char_vec))
factor_to_numeric <- as.numeric(pull(df, fact_vec))

logical_to_numeric 
character_to_numeric
factor_to_numeric
```

When we apply as.numeric() to different variable types:

+  The logical vector converts successfully: TRUE becomes 1 and FALSE becomes 0
+  The character vector fails to convert meaningfully, resulting in NA values for all elements
+  The factor vector converts to the underlying integer codes representing the factor levels

This helps explain why we can calculate the mean of the logical vector (as it gets implicitly converted to 1s and 0s) but not the character or factor vectors. R can only calculate means for numeric data or data that can be meaningfully converted to numeric values.
