---
title: "p8105_hw1_fx2212"
author: "Fenglin Xie"
date: "2025-09-15"
output: html_document
---

# The introduction of the dataset
```{r}
library(moderndive)
data("early_january_weather")
```
This dataset has `r ncol(early_january_weather)` variables and `r nrow(early_january_weather)` observations, including varibles as follows:`r names(early_january_weather)`. Besides, the mean temperature is `r mean(early_january_weather$temp)`.

# Make a scatterplot
```{r}
library(tidyverse)
temp_vs_time_plot <- ggplot(early_january_weather, 
                            aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time in Early January",
       x = "Time",
       y = "Temperature (°F)",
       color = "Humidity (%)") +
  theme_minimal()

print(temp_vs_time_plot)
```

From this plot, we can see the two points: first, humidity levels show a relationship with temperature patterns, with higher humidity values often coinciding with periods of elevated temperature. Second, there appears to be a slight overall increase in temperature values from January 7 to January 14

# Export the scatterplot to project directory
```{r}
ggsave("temp_vs_time_plot.png")
```

